Localisation and Sensor Privacy Using the Extended Information Filter and Private Linear-Combination Aggregation\thanksref{footnoteinfo}

### 2020-10-08 main.tex
First draft for whole introduction and literature review is there - rearranged a bit, section summary and notation now come before lit review - lit review is its own section

\section{Introduction}
Localisation methods in distributed sensor environments have long been an active topic of research []. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localisation methods, from range-free and range-based signal strength measurements [], echo-based range estimates [] and magnetic field measurements [] to the TDOA and TOA pseudorange multilateration methods []. Range-based localisation methods, including the use of radar measurments, GPS clocks and signal strength measurmeents, have found particularly large application due to their robustness and accuracy []. These methods use distance measurements, or computed estimates, to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location and observations centrally []. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicous gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made these requirements of information security and user privacy more apparent than before [14,15]. For example, a scenario of aircraft localisation in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Alternatively, autonomous vehicle state estimation may benefit from neighbouring vehicles' information while no participant wishes to share information about their hardware. In addition, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations. All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data.

Of the methods used to preserve privacy during distributed signal processing, it is common for the information exchanging portion of the algorithm to be reduced to a form of aggregation of involved parties' data. (For example, computing the mean value of sensor meausurements.) The most common methods for ensuring privacy during aggregation have been described below.
\begin{description}
    \item[Differential Privacy] These techniques formalise preserving the privacy of individual datum within an aggregation operation by introducing known distributions of noise to the resulting aggregation []. In the cases with no trusted aggregator, local differential privacy is achieved by introducing noise to individual inputs [norm,phe ones]. Although differential privacy has strong computational benefits due to its implementational simplicity, it provides relatively weak security guarantees for multi-party networks, with no guarantees in the presence of multiple colluding corrupted parties. In addition, noisy aggregation results hinder it use in scenarios where high precision is a desired property such as aircraft localisation.
    \item[Aggregator Oblivious Encryption] This method of encryption formalises the privacy of \textit{all} aggregation data from honest parties []. Encryptions of aggregation data are proven to be computationally indistingishable and guarantee that only their sum can be learned by an aggregator, whether or not they are trusted. The originally proposed method [] is restiricted to a small input domain which has been expandad and computationally simplified in later works [joye-lib,student-one]. The aggregation operation in aggregator obliviousness is, however, strictly limited to the sum operation and restricts its applicabilty to simple processing tasks such as smart-meter mean calculation [].
\end{description}
In addition to the private aggregation of data, it may also be the case that private data needs to be modified by an external party before being returned, as with privacy-preserving cloud computing. These methods have been broadly grouped into two domains and described below
\begin{description}
    \item[Fully Homomorphic Encryption] These schemes allow all algebraic computations to be performed on encrypted data without the computing party learning any information about inputs, intermediate values or results []. Although theoretically ideal for many privacy-preserving tasks, current implementations are still computationally impractical for real-time or large scale tasks [].
    \item[Partially Homomorphic Encryption] A simplification of fully homomorphic schemes, these schemes provide only a subset, typically one, of algebraic operations on encrypted data []. The reduced computational requirements and implementation simplicty have made partially homomorphic encryption schemes commonly used in privacy-preserving signal processing tasks, including private matrix multilpication [], set intersection computation [] and control input aggregation []. However, these works are relatively restricted in application due to the limited operations provided.
\end{description}

% TODO Never talked about FE and it's distributed extensio MCFE. General solutions still too computationally infeasible, while very specific functions do exists - dot product
% "several novel methods for privacy preserving computations ...
% which have as of yet found little application in estimation and localisation tasks

Several of the aforementioned encryption schemes have found uses in privacy-preserving localisation, estimation, and control.

% --

% Introduce localisation, filtering and the need for privacy.

% Examples of environments where privacy is relevant and concrete examples where lack of privacy could have large costs

% Methods for introducing security and privacy include differential privacy methods and encryption methods.

% Differential privacy involves using statistical noise as security to make individual users' information cannot be deduced. Often requires a trusted aggregator, although secure aggregation methods exist. always requires noising result such that the outcome is not exact (a problem in localisation).

% Encryption schemes involve formal indistinguishability proofs typically over bits or integers. They rely on computationally hard problems involving security parameters of a sufficiently large size; therefore the additional computational requirements of using encryption schemes should be pointed out and what this means in a real-time distributed sensor system. Continuing, explain public-key cryptography applicability to distributed systems; difference to symmetric schemes. Homomorphic encryption power and use case. Why FHE isn't used often, why additive partially homomorphic encryption is.

% Advancements in function providing encryption schemes such as homomorphic encryption have also led to several other types of schemes which have found uses in signal processing. Private aggregation schemes allow the secure computation of the sum of encrypted values originating from different parties, leaking only the final result. When considering such multi-party encryption protocols, formal security definitions must now also incorporate the added dangers of colluding malicious parties, and lead to new notions of security. For example Aggregator Obliviousness (AO) is typically proven for private aggregation schemes, while alternatives such as Private Weighted Secure Aggregator Obliviousness (pWSAO) exist for other specific use-cases.

% Another example of function providing encryption, and a generalisation of private aggregation, is called functional encryption (FE) and its distributed extension, multi-client functional encryption (MCFE), which allow the unencrypted result of an arbitrary function to be computed from encrypted inputs. General FE and MCFE are known to be quite computationally expensive (from meeting with ITI and student Johannes - need ref.) but alternatives providing only a subset of possibly computable function exist; for example, inner product encryption.

% Several of the aforementioned encryption schemes have found uses in secure localisation, estimation, and control.

% 
% ##       #### ######## 
% ##        ##     ##    
% ##        ##     ##    
% ##        ##     ##    
% ##        ##     ##    
% ##        ##     ##    
% ######## ####    ##    
%

\subsection{Relevant Literature on Encrypted Localisation and Estimation}

Model-free localisation using homomorphic encryption examples include polygon thing, WSN examples which protect against adversaries but in the case of the WSN paper. don't preserve anchor privacy. Importantly, model-based filtering and localisation provide more accurate estimates and these are not applicable there.

Model-based estimation examples include Aristov paper (which requires a linear model, and a hierarchy of sensors), Farokhi paper (which requires the controller compute entirely in encrypted space and send input back to actuator - supporting only the cloud-as-a-service type architectures) and Alexandru paper (which implements a distributed control environment but requires a constant gain matrix K)

pWSAO achieved in Alexandru weighted aggregation, but requires redistributing keys at every timestep resulting in a costly operation, and a complicated communication protocol.

In addition to applying suitable encryption schemes to signal processing tasks, care must be taken when converting sensor output into an encryptable homomorphic format. As is the case with our proposed localisation method, real number sensor output does not trivially encode to integers such that the homomorphic properties provided by an additive encryption scheme over integers keep the underlying real numbers consistent. Methods for handling the encoding of real numbers such that they can be used in homomorphic encryption exist. Google bignum adds power but risks overflow and leaks exponents, Farokhi leaks no information but allows only a single multiplication (extendable to more but each further multiplication limits the real number size and increases the risk of overflow).

Briefly describe navigator scenario and our contributions

%Section Summary
In Section \ref{sec:problem_statement} we introduce the localisation problem and the restrictions on involved parties. Section \ref{sec:crypto_prelim} summarises cryptographic preliminaries and Sections \ref{sec:lcao_definition} and \ref{sec:our_scheme} introduce a novel security notion of linear-combination aggregation obliviousness and an encryption scheme which satisfies it. Section \ref{sec:loc_prelim} summarises localisation preliminaries, before we introduce our private localisation method in Section \ref{sec:private_localisation}, using the introduced encryption scheme. In Section \ref{sec:sim_and_results} we give and analyse simulation results, before concluding our work in Section \ref{sec:conclusion}.

% 
% ##    ##  #######  ########    ###    ######## ####  #######  ##    ## 
% ###   ## ##     ##    ##      ## ##      ##     ##  ##     ## ###   ## 
% ####  ## ##     ##    ##     ##   ##     ##     ##  ##     ## ####  ## 
% ## ## ## ##     ##    ##    ##     ##    ##     ##  ##     ## ## ## ## 
% ##  #### ##     ##    ##    #########    ##     ##  ##     ## ##  #### 
% ##   ### ##     ##    ##    ##     ##    ##     ##  ##     ## ##   ### 
% ##    ##  #######     ##    ##     ##    ##    ####  #######  ##    ## 
%

@@

>>>>

\section{Introduction}
Localisation methods in distributed sensor environments have long been an active topic of research []. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localisation methods, from range-free and range-based signal strength measurements [], echo-based range estimates [] and magnetic field measurements [] to the TDOA and TOA pseudo-range multilateration methods []. Range-based localisation methods, including the use of radar measurements, GPS clocks and signal strength measurements, have found particularly large application due to their robustness and accuracy []. These methods use distance measurements, or computed estimates, to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location and observations centrally []. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicious gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made these requirements of information security and user privacy more apparent than before [14,15]. For example, a scenario of aircraft localisation in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Alternatively, autonomous vehicle state estimation may benefit from neighbouring vehicles' information while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations. All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows private aggregation of weighted sums, before we develop a localisation filter using this scheme such that sensor and navigator privacy is preserved.

% TODO update the below
In Section \ref{sec:problem_statement} we introduce the localisation problem and the restrictions on involved parties. Section \ref{sec:crypto_prelim} summarises cryptographic preliminaries and Sections \ref{sec:lcao_definition} and \ref{sec:our_scheme} introduce a novel security notion of linear-combination aggregation obliviousness and an encryption scheme which satisfies it. Section \ref{sec:loc_prelim} summarises localisation preliminaries, before we introduce our private localisation method in Section \ref{sec:private_localisation}, using the introduced encryption scheme. In Section \ref{sec:sim_and_results} we give and analyse simulation results, before concluding our work in Section \ref{sec:conclusion}.

@@

\section{Literature Overview}
In this section we will discuss methods for providing security guarantees in signal processing tasks, and some applications to localisation and estimation problems.

\subsection{Encryption Schemes in Signal Processing}
Of the methods used to preserve privacy during distributed signal processing, it is common for the information exchanging portion of the algorithm to be reduced to a form of aggregation of involved parties' data []. The most common methods for ensuring privacy during aggregation have been described below.
\begin{description}
    \item[Differential Privacy] These techniques formalise preserving the privacy of individual datum within an aggregation operation by introducing known distributions of noise to the resulting aggregation []. In the cases with no trusted aggregator, local differential privacy is achieved by introducing noise to individual inputs [norm, PHE ones]. Although differential privacy has strong computational benefits due to its implementational simplicity, it provides relatively weak security guarantees for multi-party networks, with no guarantees in the presence of multiple colluding corrupted parties. Additionally, noisy aggregation results hinder its use in scenarios where high precision is a desired property such as aircraft localisation.
    \item[Aggregator Oblivious Encryption] This method of encryption formalises the privacy of \textit{all} aggregation data from honest parties []. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their sum can be learned by an aggregator, whether or not they are trusted. The originally proposed method [] is restricted to a small input domain which has been expanded and computationally simplified in later works [joye-lib,student-one]. The aggregation operation in aggregator obliviousness is, however, strictly limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter mean calculation [].
\end{description}
In addition to the private aggregation of data, it may also be the case that private data needs to be modified by an external party before being returned, as with privacy-preserving cloud computing. These methods have been broadly grouped into two domains and described below.
\begin{description}
    \item[Fully Homomorphic Encryption] These schemes allow all algebraic computations to be performed on encrypted data without the computing party learning any information about inputs, intermediate values or results []. Although theoretically ideal for many privacy-preserving tasks, current implementations are still computationally impractical for real-time or large scale tasks [].
    \item[Partially Homomorphic Encryption] A simplification of fully homomorphic schemes, these schemes provide only a subset, typically one, of algebraic operations on encrypted data []. The reduced computational requirements and implementation simplicity have made partially homomorphic encryption schemes commonly used in privacy-preserving signal processing tasks, including private matrix multiplication [], set intersection computation [] and control input aggregation []. However, these works are relatively restricted in application due to the limited operations provided.
\end{description}
Lastly, recent developments in encryption schemes have provided some novel methods also suitable for privacy-preserving computation in distributed environments.
\begin{description}
    \item[Multi-Client Function Encryption] This type of encryption allows the computing of a specific function given encrypted inputs []. It is suitable for privacy-preserving distributed data as encryptions must be combined before a result is obtained, and can be considered a generalisation of the aggregation schemes described earlier. While such schemes that can compute arbitrary functions are still too computationally expensive for practical use [], partial schemes supporting only some functions, such as computing vector dot products [], exist, but have yet to find applications in more complicated signal processing tasks.
\end{description}
Of the aforementioned encryption schemes, some have found uses in privacy-preserving localisation and estimation tasks and we summarise these in the next section.

\subsection{Encrypted Localisation and Estimation}
Localisation methods exist for a variety of purposes requiring different levels of accuracy and supporting various forms of sensors []. We aim to provide an accurate localisation filter in the presence of range-only sensors.

Relevant estimation methods include both model-based Bayesian estimation methods which make assumptions about target trajectories to form their location estimates, typically based on the Kalman Filter [], and model-free methods where location estimates are made independently of each other [proloc,serloc]. The work in [proloc] uses partially homomorphic encryption to compute model-free location estimates while preserving the privacy of sensor measurements and locations. The work proposes two approaches, using either polygon intersections suffering from geometric dilution of precision, or alternating projection requiring repeated interactive comparison protocols. Neither method considers measurement uncertainty, and no estimate errors are provided.

Model-based approaches can achieve better accuracy due to their exploitation of process dynamic knowledge. In [aristov], partially homomorphic encryption is used to encrypt measurement information and produce location estimates at a navigator. While producing estimate errors and requiring only uni-directional communication, this method exclusively supports linear measurement models and is therefore not suitable for range-only measurements. Model-based encrypted control [farohki summary] also introduces some privacy-preserving methods relevant to estimation. [alexandru] proposes secure distributed control input aggregation using a novel weighted sum aggregation scheme, but requires repeated key distribution among all involved parties. In [farohki], a secure cloud-based control aggregator using partially homomorphic encryption is proposed, but is suitable only for untrusted service-providing cloud architectures.

While we primarily focus on applying encryption schemes to localisation algorithms, homomorphic encryption schemes also require suitable number encoding when used with real numbers. As is the case with localisation sensors, real number measurements need to be encoded as integers with operations consistent under scheme-provided homomorphic operations. The Google BigNum library [google bignum] provides an encoding which supports both addition and unencrypted multiplication when used with an additively homomorphic encryption scheme. It, however, leaks the exponent information of encrypted numbers. In [farohki], a simpler alternative for encoding is proposed, which leaks no information about encrypted numbers, but allows only a single unencrypted multiplication between encodings.
